---
title: "Package Details"
author: "Satyajit Sarangdhar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{"Package Details"}
  %\VignetteEncoding{UTF-8}
---

The vignetter aims to elaborate of this package.It uses the `US National Highway Traffic Safety Administration's Fatality Analysis Reporting System` data and has functions which help in summarizing and mapping the data onto US states.



```{r knitr_config,cache=FALSE,include=FALSE}
library(dplyr)
library(knitr)
#opts_knit$set(root.dir = '../data')
```

## Package Info

The package has 5 functions.

1. fars_read
2. make_filename
3. fars_read_years
4. fars_summarize_years
5. fars_map_state

```{r sourcing_code_file,messages=FALSE,warnings=FALSE,echo=FALSE,include=TRUE}
source('../R/fars_functions.R')
```


## Read csv files with fars_read()
This function reads data from a csv file or a zipped file with csv data,if it exists.
If the file doesnt exist it gives an error message saying the `file filename does not exist`.

The function takes in a string containing Path to the file if file is not stored in current directory 
and filename if the csv file is stored in current directory.

Below is output for first 5*10 matrix subset of the data for the year 2013 read using the csv file.

```{r,fars_read}
file_2013<-fars_read('accident_2013.csv.bz2')
knitr::kable(head(file_2013)[,1:10])
```


## Make data file names with make_filename()

This function reads a year `character` or `integer` and creates a filename for a csv type zipped file containing the accident  data.

The output is a string in the following format `accident_year.csv.bz`

```{r,make_file}
make_filename(2013)
```


## Read data for set of years using fars_read_years()

This function reads a vector  `character` or `integer` of years and returns a list of dataframes containing the
month of accident for each accident in the given year.If for any reason the program is unable to read
data for a paricular year an warning  `invalid year: year` is displayed and the `NULL` is returned for the given year.

The function takes in a vector of years.It first uses the make_filename function to create a filename.
The looks for a file with said filename and reads data from it using fars_read, to retain only the month and year column before moving onto next year.

As described if the program is unable to read data for particular year it throws an warning message and returns a `NULL` for the given year. 

```{r,fars_read_years,messages=FALSE,include=TRUE}
data_years<-fars_read_years(2013:2015)
str(data_years)
```


## Summarize monthly accident data for set of years using fars_summarize_years()

This function reads a vector `character`or `integer` of years and returns a dataframe summarizing total number of accidents for each month for given year in set of years. A warning message will be thrown if the program cannot find data file for a particular year.

The function takes in a vector of years.Uses fars_read_year to create a list of dataframes containing accident month and year data.Groups data by year and month and then summarizes it using dplyr functions and then creates dataframe with months as rows and a column for each year using spread function.

A warning message will be thrown if the program cannotfind data file for a particular year `invalid year: year`. 
This behavior is derived from the fars_read_years function

As described if the program is unable to read data for particular year it throws an warning message and returns a `NULL` for the given year. 

```{r,fars_summarize_years,messages=FALSE,include=TRUE}
summarize_years<-fars_summarize_years(2013:2015)
summarize_years
```

## Map accident data onto states using fars_map_state()

The function reads a state number and the year and plots the accident areas in the state on a map.
If the state number entered doesnt belong to a valid state in US an error is thrown saying `invalid STATE number: state num`  and program stops execution.

The function takes in state number and year, then using both and make_filename and fars_read reads in
data for particular year and then filters it out for a particular state num using dply's filter function.If a invalid state num is entered the function throws an error and stops the program.

This data is then checked for any spurious values of longitude and latitude which are replaced by NAs.
Then using the maps package's map function the US state area is plotted and using the graphics package's
points function the accident areas marked on the map.

The program will also stop execution if a data file for the particular year is not found in the current directory.This behavior is derived from the fars_read function of the package.
If there is no accident given year in the particular state then the message `no accidents to plot` is displayed.

```{r,fars_map_state,messages=FALSE}

fars_map_state(22,2013)

```







